---
title: "p8105_hw2_jv3014"
author: "Jia Varkey"
date: "2026-10-06"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1


```{r}
transit_df = read_csv("./HW2Data/NYC_Transit.csv") |>
  janitor::clean_names() |>
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada) |>
  mutate(
    entry = case_match(entry, 
                       "YES" ~ TRUE,
                       "NO" ~ FALSE)
  ) 


```


```{r}
distinct_station = 
  transit_df |>
  distinct(line, station_name)

ada_compliant =
  transit_df |>
  filter(ada==TRUE)


no_vending_entrance =
  transit_df |>
  filter(vending == "NO", entry == TRUE)

total_no_vending =
  transit_df |>
  filter(vending == "NO")

entrance_vend_prop = nrow(no_vending_entrance) / nrow(total_no_vending) * 100
```
There are `r nrow(distinct_station)` distinct stations, with `r nrow(ada_compliant)` ADA compliant. `entrance_vend_prop`% of the stations with no vending allow entrance. 

```{r}

route_combined = 
  transit_df |>
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(
    cols = starts_with("route"),
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )


distinct_A =
  route_combined |>
  filter(route_name == "A")

A_ada =
  distinct_A |>
  filter(ada==TRUE)
```
There are `r nrow(distinct_A)` distinct stations that serve the A train. Of those stations, `r nrow(A_ada)` are ADA compliant. 

## Problem 2

```{r}
trash_df = 
  readxl::read_excel("./HW2Data/Trashwheel.xlsx", sheet = "Mr. Trash Wheel") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  select(-x15, -x16) |>
  mutate(
    sports_balls = 
      as.integer(
      round(sports_balls)),
    wheel = "Mr. Trash Wheel",
    year = as.numeric(year)
  ) 

prof_df =
  readxl::read_excel("./HW2Data/Trashwheel.xlsx", sheet = "Professor Trash Wheel") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(wheel = "Professor Trash Wheel",
         plastic_bottles = as.numeric(plastic_bottles))

gwynnda_df =
  readxl::read_excel("./HW2Data/Trashwheel.xlsx", sheet = "Gwynnda the Good Wheel of the W") |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(wheel = "Gwynnda")

# Combine datasets

trashwheel_df = bind_rows(trash_df, prof_df, gwynnda_df) |>
  janitor::clean_names()
```

```{r}
prof_total_weight =
  trashwheel_df |>
  filter(wheel == "Professor Trash Wheel") |>
  summarize(total_weight = sum(weight_tons, na.rm = TRUE)) |>
  pull(total_weight)

gwynnda_cig = 
  trashwheel_df |>
  filter(wheel == "Gwynnda",
         year == 2022, 
         month == "June") |>
  summarize(total_cigarettes = sum(cigarette_butts, na.rm = TRUE)) |>
  pull(total_cigarettes)
```

The combined trash wheel dataset contains a total of `r nrow(trashwheel_df)` observations tracking data across three trash wheels. Key variables include the total weight of the dumpster in tons as `weight_tons`, `volume_cubic_yards`, and different types of trash 
collected such as `plastic_bottles`, `wrappers`, and `sports_balls`.

Based on the available data, Professor Trash Wheel collected a total trash weight of `r prof_total_weight` tons and the total number of cigarette butts that Gwynnda collected in June 2022 is `r gwynnda_cig`. 

rename(
    ID = x1,
    age = age_at_the_study_baseline,
    sex = x1_male_0_female,
    apoe4 = x1_apoe4_carrier_0_apoe4_non_carrier,
    age_onset = age_at_the_onset_of_mci_missing_if_a_subject_remains_mci_free_during_the_follow_up_period
  ) |>

## Problem 3

```{r}

mci_baseline = 
  read_csv("data_mci/MCI_baseline.csv", skip = 1, na = c("", "NA", "Na", "."))|>
  janitor::clean_names() |>
  
  mutate(
    sex = factor(
      sex, 
      levels = c(0,1),
      labels = c("Female", "Male")
      ),
    
    apoe4 = factor(
      apoe4,
      levels = c(0,1),
      labels = c("Non-carrier", "Carrier")
    )
  ) 

```



```{r}
avg_baseline_age = 
  mci_baseline |>
  summarize(mean_age = mean(current_age, na.rm = TRUE)) |>
  pull(mean_age)


women_prop =
  mci_baseline |>
  filter(sex == "Female") |>
  summarise(mean(apoe4 == "Carrier"))
  
number_participants = nrow(mci_baseline) - 1

developed_mci = 
  mci_baseline |>
  filter(!is.na(age_at_onset)) |>
  nrow()
```
  
The MCI baseline data set shows the onset and development of MCI. `r number_participants` people were recruited and from this sample, `r developed_mci ` developed MCI. The average baseline age is `r avg_baseline_age` and the proportion of women in the study who are carriers is `r women_prop`. 

Import amyloid data 
```{r}
mci_amyloid =
  read_csv("data_mci/mci_amyloid.csv", skip = 1, na = c("", "NA", "Na", ".")) |>
  janitor::clean_names() |>
  rename(
    id = study_id,
  )
```

```{r}
anti_join(mci_baseline, mci_amyloid, by = "id") 
anti_join(mci_amyloid, mci_baseline, by = "id")
```

8 participants in the baseline data do not have biomarker data. Whereas 16 of those in the amyloid data did not participate in the baseline data. 


```{r}
combined_mci = inner_join(mci_baseline, mci_amyloid, by = "id")
write_csv(combined_mci,"data_mci/combined_mci.csv)")
```

